---
title: "coffee"
author: "Shuqi Cao"
output:
  pdf_document:
          latex_engine: pdflatex
          number_sections: yes
fig_caption: yes
---


```{r loadpackages, echo=FALSE, eval=TRUE, warning=FALSE, message=FALSE}
library(tidyverse)
library(gridExtra)
library(kableExtra)
library(jtools)
library(sjPlot)
```


# introduction

## Research question

What influence do different features of coffee have on whether the quality of a batch of coffee is classified as good or poor?

## Data description

The dataset is collected from the Coffee Quality Database (CQD) of Coffee Quality Institute.  As a non-profit organisation, the institute aims to improve the quality of coffee and the lives of farmers who produce the beans. The dataset contains information on features of coffee and its production, including an overall quality score.  
 
**Response variable**

* `country_of_origin` – Country where the coffee bean originates from. 

**Explanatory variables**

* `aroma` – Aroma grade (ranging from 0-10)
* `flavor` – Flavour grade (ranging from 0-10)
* `acidity` – Acidity grade (ranging from 0-10) 
* `category_two_defects` – Count of category 2 type defects in the batch of coffee beans tested.
* `altitiude_mean_meters` – Mean altitude of the growers farm (in metres)
* `harvested` – Year the batch was harvested
* `Qualityclass` – Quality score for the batch (Good - >= 82.5, Poor - < 82.5). Note: 82.5 was selected as the cut off as this is the median score for all the batches tested. 


# Data summarisation

We can load in our data set and see what it looks like by using the glimpse function.

```{r data, echo = FALSE, eval = TRUE, warning = FALSE}
coffee <- read.csv("dataset13.csv")
summary(coffee)
str(coffee)
table(coffee$Qualityclass)
```

```{r clean, echo = FALSE, eval = TRUE, warning = FALSE}
data <- coffee %>% filter(#is.na(altitude_mean_meters)==FALSE, 
                          is.na(harvested)==FALSE, 
                          altitude_mean_meters<9000,
                          aroma!=0) #aroma*flavor*acidity!=0
```

```{r boxp, echo = FALSE, eval = TRUE, warning = FALSE, fig.pos = "H"}
p1 <- ggplot(data = data, aes(x = Qualityclass, y = aroma, fill = Qualityclass)) +
  geom_boxplot() +
  labs(x = "quality class", y = "aroma") + 
  theme(legend.position = "none")
p2 <- ggplot(data = data, aes(x = Qualityclass, y = flavor, fill = Qualityclass)) +
  geom_boxplot() +
  labs(x = "quality class", y = "flavor") + 
  theme(legend.position = "none")
p3 <- ggplot(data = data, aes(x = Qualityclass, y = acidity, fill = Qualityclass)) +
  geom_boxplot() +
  labs(x = "quality class", y = "acidity") + 
  theme(legend.position = "none")
p4 <- ggplot(data = data, aes(x = Qualityclass, y = category_two_defects, fill = Qualityclass)) +
  geom_boxplot() +
  labs(x = "quality class", y = "category_two_defects") + 
  theme(legend.position = "none")
p5 <- ggplot(data = data, aes(x = Qualityclass, y = altitude_mean_meters, fill = Qualityclass)) +
  geom_boxplot() +
  labs(x = "quality class", y = "altitude_mean_meters") + 
  theme(legend.position = "none")
grid.arrange(p1, p2, p3, p4, p5, ncol = 2)
```

```{r barp, echo = FALSE, eval = TRUE, warning = FALSE, fig.pos = "H"}
ggplot(data = data, aes(x= Qualityclass, y = ..prop.., group = harvested, fill = harvested)) +
  geom_bar(position="dodge", stat="count") +
  labs(y = "proportion")
```

```{r level, echo = FALSE, eval = TRUE, warning = FALSE}
data$Qualityclass <- factor(data$Qualityclass, levels = c("Poor", "Good"))
data$harvested <- factor(data$harvested, levels = c("2010", "2011", "2012", "2013", "2014", "2015", "2016", "2017", "2018"))
str(data)
dat1 <- data[,-1]
```


# Methods

## Log-odds

model1

```{r model1, echo = FALSE, eval = TRUE, warning = FALSE, fig.pos = "H"}
mod1 <- glm(Qualityclass ~ ., data = dat1, family = binomial(link = "logit"))
mod1 %>% summary()
```

model2

```{r model2, echo = FALSE, eval = TRUE, warning = FALSE, fig.pos = "H"}
dat2 <- dat1[,-6]
mod2 <- glm(Qualityclass ~ ., data = dat2, family = binomial(link = "logit"))
mod2 %>% summary()
```

model3

```{r model3, echo = FALSE, eval = TRUE, warning = FALSE, fig.pos = "H"}
dat3 <- dat2[,-4]
mod3 <- glm(Qualityclass ~ ., data = dat3, family = binomial(link = "logit"))
mod3 %>% summary()
```

```{r lodds3, echo = FALSE, eval = TRUE, warning = FALSE, fig.pos = "H"}
confint(mod3) %>%
  kable()%>%
  kable_styling(font_size=10, latex_options="HOLD_position")
plot_model(mod3, show.values = TRUE, transform = NULL,
           title = "Log-Odds", show.p = TRUE)
```

optional model4 (considered 95% CI)

```{r model4, echo = FALSE, eval = TRUE, warning = FALSE, fig.pos = "H"}
dat4 <- dat3[,-4]
mod4 <- glm(Qualityclass ~ ., data = dat4, family = binomial(link = "logit"))
mod4 %>% summary()
```

```{r lodds4, echo = FALSE, eval = TRUE, warning = FALSE, fig.pos = "H"}
confint(mod4) %>%
  kable()%>%
  kable_styling(font_size=10, latex_options="HOLD_position")
plot_model(mod4, show.values = TRUE, transform = NULL,
           title = "Log-Odds (Good instructor)", show.p = TRUE)
```


## Model selection

```{r comp, echo = FALSE, eval = TRUE, warning = FALSE}
model.comp1 <- glance(mod1)
model.comp2 <- glance(mod2)
model.comp3 <- glance(mod3)
model.comp4 <- glance(mod4)
Models <- c("GLM1","GLM2","GLM3","GLM4")
bind_rows(model.comp1, model.comp2, model.comp3, model.comp4, .id="Model") %>%
  select(Model,AIC,BIC) %>%
  mutate(Model=Models) %>%
  kable(
    digits = 3, align = "c", col.names = c("model","AIC","BIC"),
    caption = "\\label{tab:select} Model comparison values for different models",
  ) %>%
  kable_styling(font_size=10, latex_options="HOLD_position")
```

deviance (? whether use or not)

```{r anova, echo = FALSE, eval = TRUE, warning = FALSE, fig.pos = "H"}
mod_base <- glm(Qualityclass ~ 1, data = dat3, family = binomial(link = "logit"))
anova(mod_base,mod3,test="Chisq") 
```

## Odds

```{r odds, echo = FALSE, eval = TRUE, warning = FALSE, fig.pos = "H"}
mod3 %>%
  coef() %>%
  exp() %>%
  kable(digits = 3, align = "c", col.names = "Odds", caption = "Odds-ratio") %>%
  kable_styling(font_size=10, latex_options="HOLD_position")
plot_model(mod3, show.values = TRUE, #axis.lim = c(1,1.5),
           title = "Odds (Good instructor)", show.p = TRUE)
```


## Probabilities

```{r prob, echo = FALSE, eval = TRUE, warning = FALSE, fig.pos = "H"}
dat3 <- dat3 %>% mutate(probs.good = fitted(mod3))
p <- plot_model(mod3, type = "pred", title = "")
plot_grid(p)
```

